---
title: "Models_and_Post-Hocs"
format: html
editor: visual
---

# Read data, factorize variables, and specify contrasts

```{r}
library(readr)
library(dplyr)
library(bestNormalize)
library(emmeans)
library(lme4)
library(readxl)
library(effectsize)

df = read_csv("dataset.csv")
df$automation = factor(df$automation)
df$nback = factor(df$nback)
df$runNumber = factor(df$runNumber)
contrasts(df$nback) <- contr.sum
contrasts(df$automation) <- contr.sum
contrasts(df$runNumber) <- contr.sum
```
All subsequent analyses were organized first by metric and then by hazard type. Each workflow comprised data filtering, model specification and fitting, and post hoc analyses. 

# Reaction Time

## Left Pedestrian

```{r}

df2 = subset(df,eventType=="LEFT_PEDESTRIAN")

LMM<-lmerTest::lmer(rt~ (automation+nback)^2+Age+(1| subNumber), data=df2)

anova(LMM)

emm <- emmeans(LMM, ~ nback)

contrast(emm, "pairwise", adjust = "bonferroni") |>
  summary(infer = TRUE, infer.df = TRUE)

eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))

```

## Right Pedestrian

```{r}

df2 = subset(df,eventType=="RIGHT_PEDESTRIAN")

LMM<-lmerTest::lmer(rt~ automation*nback+Age+(1 |subNumber), data=df2)

anova(LMM)

emm1 <- emmeans(LMM, ~ nback|automation)

contrast(emm1, "pairwise", adjust = "bonferroni") |>
  summary(infer = TRUE, infer.df = TRUE)

eff_size(emm1, sigma = sigma(LMM), edf = df.residual(LMM))

emm2 <- emmeans(LMM, ~ automation|nback)

contrast(emm2, "pairwise", adjust = "bonferroni") |>
  summary(infer = TRUE, infer.df = TRUE)

eff_size(emm2, sigma = sigma(LMM), edf = df.residual(LMM))

```

## Oncoming Vehicle

```{r}

df2 = subset(df,eventType=="ONCOMING_VEHICLES")

yj_obj <- yeojohnson(df2$rt)
df2$y_yj <- yj_obj$x.t

LMM<- lmerTest::lmer (y_yj~ automation*nback+runNumber + (1 + nback || subNumber) ,data = df2)

anova(LMM)

linkfun_yj <- function(mu) as.numeric(predict(yj_obj, newdata = mu))
linkinv_yj <- function(eta) as.numeric(predict(yj_obj, newdata = eta, inverse = TRUE))

mu_eta_yj <- function(eta, eps = 1e-6) {
  (linkinv_yj(eta + eps) - linkinv_yj(eta - eps)) / (2 * eps)
}

yj_link <- list(
  linkfun = linkfun_yj,
  linkinv = linkinv_yj,
  mu.eta  = mu_eta_yj,
  valideta = function(eta) TRUE,
  name = "Yeo-Johnson (bestNormalize)"
)

emm <- emmeans(LMM, ~ runNumber)
pairs(emm, adjust = "bonferroni") |>
  summary(infer = TRUE, type="response")
eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))

emm_bt <- regrid(update(emm, tran = yj_link))
pairs(emm_bt, adjust = "bonferroni") |>
  summary(infer = TRUE, type="response")
```

```{r}
# Remove Two Trials Where Programming Errors Prevented Takeovers
df = subset(df,!((subNumber==36&runNumber==1&eventType=="RIGHT_PEDESTRIAN")|(subNumber==11&runNumber==1&eventType=="RIGHT_PEDESTRIAN")))
```

# Reaction Type

## Left Pedestrian

```{r}

df2 = subset(df,eventType=="LEFT_PEDESTRIAN")

GLMM <- glmer (steering ~ (automation+nback)^2+(1|subNumber), family=binomial
               ,data = df2)

car::Anova(GLMM)

emmeans(GLMM, ~ automation|nback) |>
  contrast("revpairwise", adjust = "bonferroni") |>
  summary(infer = TRUE,type="response")

emmeans(GLMM, ~ nback|automation) |>
  contrast("pairwise", adjust = "bonferroni") |>
  summary(infer = TRUE,type="response")
```

## Right Pedestrian

```{r}

df2 = subset(df,eventType=="RIGHT_PEDESTRIAN")

GLMM <- glmer (steering ~ (automation+nback)^2+(1+nback||subNumber), family=binomial
               ,data = df2)

car::Anova(GLMM)

```

## Oncoming Vehicles

```{r}

df2 = subset(df,eventType=="ONCOMING_VEHICLES")

GLMM <- glmer (steering ~ (automation+nback)^2+(1|subNumber), family=binomial
               ,data = df2)

car::Anova(GLMM)

```

# Vehicular Jerk

## Left Pedestrian

```{r}

df2 = subset(df,eventType=="LEFT_PEDESTRIAN")

LMM<-lmerTest::lmer(jerk~ (automation+nback)^2+runNumber+Age+nback:runNumber+(1+automation||subNumber), data=df2)

anova(LMM)

emm <- emmeans(LMM, ~ automation)

contrast(emm, "revpairwise", adjust = "bonferroni") |>
  summary(infer = TRUE, infer.df = TRUE)

eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))

```

## Right Pedestrian

```{r}

df2 = subset(df,eventType=="RIGHT_PEDESTRIAN")

yj_obj <- yeojohnson(df2$jerk)
df2$y_yj <- yj_obj$x.t

LMM<- lmerTest::lmer (y_yj~ automation*nback + (1| subNumber) ,data = df2)

anova(LMM)

linkfun_yj <- function(mu) as.numeric(predict(yj_obj, newdata = mu))
linkinv_yj <- function(eta) as.numeric(predict(yj_obj, newdata = eta, inverse = TRUE))

mu_eta_yj <- function(eta, eps = 1e-6) {
  (linkinv_yj(eta + eps) - linkinv_yj(eta - eps)) / (2 * eps)
}

yj_link <- list(
  linkfun = linkfun_yj,
  linkinv = linkinv_yj,
  mu.eta  = mu_eta_yj,
  valideta = function(eta) TRUE,
  name = "Yeo-Johnson (bestNormalize)"
)

emm <- emmeans(LMM, ~ automation)
contrast(emm, "revpairwise", adjust = "bonferroni") |>
  summary(infer = TRUE, infer.df = TRUE)

eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))


# emm_bt <- regrid(update(emm, tran = yj_link))
# 
# summary(emm_bt, infer = TRUE)
# 
# pairs(emm_bt, adjust = "bonferroni") |>
#   summary(infer = TRUE)

```

## Oncoming Vehicles

```{r}
df2 = subset(df,eventType=="ONCOMING_VEHICLES")

LMM<- lmerTest::lmer (jerk ~ automation*nback + Age+ automation:Age+(1| subNumber) ,data = df2)

anova(LMM)

emm <- emmeans(LMM, ~ automation)
contrast(emm, "revpairwise", adjust = "bonferroni") |>
  summary(infer = TRUE, infer.df = TRUE)

eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))


```

# Minimum TTC

## Left Pedestrian

```{r}

df2 = subset(df,eventType=="LEFT_PEDESTRIAN")

LMM= lmerTest::lmer(minTTC~ automation + nback + runNumber+ (1+automation||
    subNumber) +automation:nback + nback:runNumber,data=df2)

anova(LMM)

emmeans(LMM, ~ automation|nback) |>
  contrast("pairwise", adjust = "bonferroni") |>
  summary(infer = TRUE,infer.df=TRUE,digits = 4)

emm = emmeans(LMM, ~ automation|nback)
eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))

emmeans(LMM, ~ nback|automation) |>
  contrast("revpairwise", adjust = "bonferroni") |>
  summary(infer = TRUE,infer.df=TRUE,digits = 4)

emm = emmeans(LMM, ~ nback|automation)
eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))

emmeans(LMM, ~ nback|runNumber) |>
  contrast("pairwise", adjust = "bonferroni") |>
  summary(infer = TRUE,infer.df=TRUE,digits = 4)

emm = emmeans(LMM, ~ nback|runNumber)
eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))

emmeans(LMM, ~ runNumber|nback) |>
  contrast("pairwise", adjust = "bonferroni") |>
  summary(infer = TRUE,infer.df=TRUE,digits = 4)

emm = emmeans(LMM, ~ runNumber|nback)
eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))

```

## Right Pedestrian

```{r}

df2 = subset(df,eventType=="RIGHT_PEDESTRIAN")

LMM= lmerTest::lmer(minTTC~ automation * nback + Age+(1+automation+nback|
    subNumber),data=df2)

anova(LMM)

emmeans(LMM, ~ automation) |>
  contrast("pairwise", adjust = "bonferroni") |>
  summary(infer = TRUE)
emm = emmeans(LMM, ~ automation)
eff_size(emm, sigma = sigma(LMM), edf = df.residual(LMM))
```

## Oncoming Vehicles

```{r}
library(glmmTMB)
df2 = subset(df,eventType=="ONCOMING_VEHICLES")

GLMM <- glmmTMB::glmmTMB (minTTC ~ (automation+nback)^2+(1+automation+nback||subNumber), family = tweedie(link = "log"), data = df2)

car::Anova(GLMM)

emmeans(GLMM, ~ automation) |>
  contrast("pairwise", adjust = "bonferroni",type='response') |>
  summary(infer = TRUE)


```

# Crash Probability

## Left Pedestrian

```{r}

df2 = subset(df,eventType=="LEFT_PEDESTRIAN")

GLMM <- glmer (crash ~ (automation+nback)^2+(1|subNumber), family=binomial, data = df2)

car::Anova(GLMM)
```

## Right Pedestrian

```{r}
df2 = subset(df,eventType=="RIGHT_PEDESTRIAN")

GLMM <- glmer (crash ~ (automation+nback)^2+(1|subNumber), family=binomial, data = df2)

car::Anova(GLMM)
```

## Oncoming Vehicles

```{r}
df2 = subset(df,eventType=="ONCOMING_VEHICLES")

GLMM <- glmer (crash ~ (automation+nback)^2+runNumber+automation:runNumber+(1|subNumber), family=binomial, data = df2, control = glmerControl(
                 optimizer = "bobyqa"))

car::Anova(GLMM)

emmeans(GLMM, ~ automation) |>
  contrast("revpairwise", adjust = "bonferroni", type="response") |>
  summary(infer = TRUE)

```



